Approximation as Differentiation: Notes

### 2025-07-04 notes/predicate-system.tex
Reindex/direct image along a natural transformation.

Suppose $X$ a set. Define a \emph{predicate on $X$} to be a family of (proof-irrelevant) propositions. For any
function $f: X \to Y$, define $\reindex{P}{f}(x)$ iff $P(f(x))$ and $\directImage{P}{f}(y)$ iff there exists
$x$ such that $f(x) = y$ and $P(x)$. Define $\mathcal{S}$ to be the predicate system over $\Set$ which assigns
to any set $X$ the set $\Pred(X)$ of propositions on $X$ ordered by implication, with $\top, \meet, \join$ and
$\residual$ defined pointwise and universal quantification defined as $\forall_Y(P)(x)$ iff $P(x,y)$ for all
$y \in Y$.

@@

Now define the predicate system $\mathcal{S}^{\cat{C}}$ over $\Func{\cat{C}^\op}{\Set}$ which assigns to every
presheaf $F$ the poset $(\Pred(F), \sqleq)$ of predicates on $F$ ordered pointwise, and where $\top, \meet,
\join, \residual$ and $\forall$ in $\Pred(F)$ are all given pointwise. For every natural transformation
$\alpha: F \to G$, the pullback $\reindex{-}{\alpha}: \Pred(G) \to \Pred(F)$ ...

>>>>

Suppose $X$ a set. Define a \emph{predicate on $X$} to be a family of (proof-irrelevant) propositions for
every $x \in X$. For any function $f: X \to Y$, define $\reindex{P}{f}(x)$ iff $P(f(x))$ and
$\directImage{P}{f}(y)$ iff there exists $x$ such that $f(x) = y$ and $P(x)$. Define $\mathcal{S}$ to be the
predicate system over $\Set$ which assigns to any set $X$ the set $\Pred(X)$ of propositions on $X$ ordered by
implication, with $\top, \meet, \join$ and $\residual$ defined pointwise and universal quantification defined
as $\forall_Y(P)(x)$ iff $P(x,y)$ for all $y \in Y$.

@@

Now define the predicate system $\mathcal{S}^{\cat{C}}$ over $\Func{\cat{C}^\op}{\Set}$ which assigns to every
presheaf $F$ the poset $\Pred(F)$ of predicates on $F$ ordered pointwise, and where $\top, \meet, \join,
\residual$ and $\forall$ in $\Pred(F)$ are all given pointwise. For every natural transformation $\alpha: F
\to G$, define $\reindex{P}{\alpha}_X = \reindex{P_X}{\alpha_X}$ and $\directImage{P}{\alpha}_X =
\directImage{P_X}{\alpha_X}$.
